---
title: 'Reproducible Research: Peer Assessment 1'
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Loading and preprocessing the data

```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile = "activity.zip", mode="wb")
unzip("activity.zip")
stepdata = read.csv("activity.csv", header = TRUE)
```

## 2. What is mean total number of steps taken per day?

```{r}
library(magrittr)
library(dplyr)

databydate = stepdata %>% 
        select(date, steps) %>% 
        group_by(date) %>% 
        summarize(tsteps= sum(steps)) %>%
        na.omit() %>%
        print(n=60)

databydate = stepdata %>% 
        select(date, steps) %>% 
        group_by(date) %>% 
        summarize(tsteps= sum(steps)) %>%
        na.omit() %>%
        print(n=60)
```

## 3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(databydate$tsteps)
median(databydate$tsteps)
```

## 4. Time series plot

```{r}
library(ggplot2)

databyinterval <- stepdata %>% 
        select(interval, steps) %>% 
        na.omit() %>% 
        group_by(interval) %>% 
        summarize(tsteps= mean(steps)) 

ggplot(databyinterval, aes(x=interval, y=tsteps)) + geom_line()
```

# 5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
databyinterval[which(databyinterval$tsteps == max(databyinterval$tsteps)),]
```

# 6. Code to describe and show a strategy for imputing missing data
```{r}
# code to describe
missingVals <- sum(is.na(data))
## Warning in is.na(data): is.na() applied to non-(list or vector) of type
## 'closure'
missingVals
## [1] 0

# strategy for imputing missing data
replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))

meandata <- stepdata %>% 
        group_by(interval) %>% 
        mutate(steps= replacewithmean(steps))

meandata

```

# 7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
FullSummedDataByDay <- aggregate(meandata$steps, by=list(meandata$date), sum)

FullSummedDataByDay

names(FullSummedDataByDay)[1] ="date"
names(FullSummedDataByDay)[2] ="totalsteps"

summary(FullSummedDataByDay)

hist(FullSummedDataByDay$totalsteps, xlab = "Steps", ylab = "Frequency", main = "Total Daily Steps", breaks = 20)
```

# 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r}
meandata$date <- as.Date(meandata$date)
meandata$weekday <- weekdays(meandata$date)
meandata$weekend <- ifelse(meandata$weekday=="Saturday" | meandata$weekday=="Sunday", "Weekend", "Weekday" )

library(ggplot2)

meandataweekendweekday <- aggregate(meandata$steps , by= list(meandata$weekend, meandata$interval), na.omit(mean))

names(meandataweekendweekday) <- c("weekend", "interval", "steps")

ggplot(meandataweekendweekday, aes(x=interval, y=steps, color=weekend)) + geom_line() +
        facet_grid(weekend ~.) + xlab("Interval") + ylab("Mean of Steps") +
        ggtitle("Comparison of Average Number of Steps in Each Interval")
```

